---
title: "Using `freezr` for tracking data analyses"
author: "Eric Kernfeld"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{freezr tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


###What problem does `freezr` solve?

Data analysis projects require moderately complex organization 
(
[example 1](https://www.maxmasnick.com/analysis-org/), 
[example 2](https://r-dir.com/blog/2013/11/folder-structure-for-data-analysis.html), 
[example 3](https://medium.com/human-in-a-machine-world/folder-structure-for-data-analysis-62a84949a6ce#.tj5h1u2wp)
), which can be daunting, especially for newcomers to the field.
To make a project reproducible, the data analyst incurs even more overhead. Every time 
a result merits recording, their future self (or a colleague taking over) will need 
more than just the current graphical or text output. They'll need, at a minimum, working code and 
notes about the analysis. Better still would be to save package versions and other 
inputs or dependencies. 

Fortunately, much of this extra work can be automated. 

###`freeze` helps record analyses for later
 
First, some preliminaries: this will create some fake code that you can pretend you've written.

```{r, warning=F}
fake_proj = file.path( getwd(), "fake_project" )
dir.create(fake_proj); setwd(fake_proj)
cat("my_fun = function(x) plot(x)", file = file.path( fake_proj, "my_functions.R" ) )
cat("my_fun(1:5);print('blah')", file = file.path( fake_proj, "my_script.R" ) )
dir.create(file.path( fake_proj, "results" ))
```

Now let's freeze it.

```{r, warning=F}
library(freezr)
freeze( analyses_to_run = c( "my_functions.R", "my_script.R" ),
        destination = file.path( fake_proj, "results" ) )
```
   
In this line of code, `freeze` has:

- **run your scripts**,
- **gathered the output** into a time-stamped results folder,
- **save a copy of your code** along with the results, and
- **nagged you to write down some notes** about what you did.

Now you can go add notes to the file it created, and then you're free to keep iterating on your analysis.

###Inventory

Let's move into another demo folder.

```{r, warning=F}
dir.create(path = "inventory_demo")
setwd("inventory_demo")
```

Imagine you have a bunch of crap built up, including the alphabet.
```{r, warning=F}
ugly_folders = paste0( "long_ugly_results_folder_2017JAN11", 1:2)
lapply( ugly_folders, dir.create )
write.table(letters, file.path(ugly_folders[1], "alphabet.txt"))
```

You're about to add Spanish characters, and you wanted to easily retrieve the alphabet.
So you put it in the inventory with an easy-to-remember tag and some notes to yourself.
```{r}
inventory( inv_location = getwd(),
           tag = "abc",
           filename = file.path(ugly_folders[1], "alphabet.txt"),
           extra = "Honestly this is just the alphabet.")
inventory( inv_location = getwd() )
```

Then you fetch it from the inventory and save a new version somewhere else.
```{r}
old_alpha_file = inventory( inv_location = getwd(),
                            tag = "abc")
old_alpha = read.table(old_alpha_file, stringsAsFactors = F)[[1]]
write.table(sort(c(old_alpha, "rr", "ch", "ll", "ñ")), file.path(ugly_folders[2], "alphabet.txt"))
```

You can add a new record with language awareness and delete the old one:
```{r}
inventory( inv_location = getwd(),
           tag = "abc_eng",
           filename = file.path(ugly_folders[1], "alphabet.txt"),
           extra = "This is the English alphabet.")
inventory( inv_location = getwd(),
           tag = "abc",
           delete = T )
inventory( inv_location = getwd() )
```

Finally, you add a record for the Spanish alphabet.
Using a `parent_tag`, you can include the fact that
the construction of the Spanish alphabet version
depended on the English alphabet version.

```{r}
inventory( inv_location = getwd(),
           tag = "abc_esp",
           parent_tag = "abc_eng",
           filename = file.path(ugly_folders[2], "alphabet.txt"),
           extra = "This is the Spanish alphabet.")
inventory( inv_location = getwd() )
```

Move back out of the demo folder to finish up.

```{r}
setwd(dirname(getwd()))
```

